---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = FALSE, message = FALSE, warning = FALSE,
  comment = "#>",
  fig.path = "README-",
  fig.align = "center"
)
```

# auklet: Analysis and visualization of your eBird sightings

`auklet` provides tools for analyzing and visualizing your personal [eBird](http://ebird.org) data. Your personal sightings can be downloaded as a CSV file from the [Download My Data](http://ebird.org/ebird/downloadMyData) page on the eBird website.

## Installation

Install `auklet` from GitHub using:

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("CornellLabofOrnithology/auk")
```

## Usage

All functions in `auklet` begin with `eb_` (for eBird) to aid tab completion. Import your eBird sightings data into a data frame with `eb_sightings()`:

```{r import}
library(auklet)
library(dplyr)
# load example data inclued with the package
ebird_data <- system.file("extdata/MyEBirdData.csv", package = "ebird") %>%
  eb_sightings()
ebird_data
```

In addition to true, countable species, your eBird data includes other taxa that can't be resolved to the species level (e.g. "spuhs" such as *Empidonax sp.*) or aren't countable (e.g. domestic species). Other taxa in your data may be reported at a level below species (e.g. subspecies or recognizable forms). `eb_sightings()` contains additional fields to help in resolving these issues. `species_code`, `species_common`, and `species_scientific` are `NA` for taxa that aren't resolvable to countable species, and give the corresponding species for taxa reported below species.

Once your eBird data are imported, you can begin summarizing and visualizing them. For example, use `eb_daylist()` to creat daily life lists, i.e. a data frame of species seen on each day of the year.

```{r day-list}
day_lists <- eb_daylist(ebird_data)
# species seen on feb 14
filter(day_lists, month == 2, day == 14) %>% 
  select(month, day, species_common)
```

These day lists can be summarized to daily counts with `summary()` or visualized with `plot()`.

```{r summ-plot, fig.width = 10, fig.height = 4, fig.retina = TRUE}
summary(day_lists) %>% 
  head()
plot(day_lists)
```

