---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = FALSE, message = FALSE, warning = FALSE,
  comment = "#>",
  fig.path = "README-",
  fig.align = "center"
)
```

# auklet: Analysis and visualization of your eBird sightings

`auklet` provides tools for analyzing and visualizing your personal [eBird](http://ebird.org) data. Your personal sightings can be downloaded as a CSV file from the [Download My Data](http://ebird.org/ebird/downloadMyData) page on the eBird website.

## Installation

Install `auklet` from GitHub using:

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("mstrimas/auklet")
```

## Usage

All functions in `auklet` begin with `eb_` (for eBird) to aid tab completion. Import your eBird sightings data into a data frame with `eb_sightings()`:

```{r import}
library(auklet)
library(dplyr)
# load example data inclued with the package
ebird_data <- system.file("extdata/MyEBirdData.csv", package = "auklet") %>%
  eb_sightings()
```

Once your eBird data are imported, you can begin summarizing and visualizing them. The most basic functionality is generating your life list.

```{r life-list}
eb_lifelist(ebird_data) %>% 
  select(species_common, date, country) %>% 
  head()
```

Life lists can, of course, be viewed directly on the eBird website; however, other functions produce summaries or visualizations not available in eBird. For example, use `eb_lifelist_day()` to creat daily life lists, i.e. a data frame of species seen on each day of the year.

```{r day-list}
day_lists <- eb_lifelist_day(ebird_data)
# species seen on feb 14
filter(day_lists, month == 2, day == 14) %>% 
  select(month, day, species_common)
```

These day lists can be summarized to daily counts with `summary()` or visualized with `plot()`.

```{r summ-plot, fig.width = 10, fig.height = 4, fig.retina = TRUE}
summary(day_lists) %>% 
  head()
plot(day_lists)
```

## Acknowledgments

This package, and some of the specific functionality, was inspired by conversations with [Drew Weber](http://www.nemesisbird.com/), [Taylor Long](http://www.mtaylorlong.com/), and Tom Auer.
